Which ITG Stepcharts are Turniest?

### 2019-03-17 itg/3/paper.tex
finished i think

What, then, is a bracket jump (which I shall \textit{not}, henceforth, abbreviate for brevity)?
Put simply, whenever two arrow-shaped obstacles proceed simultaneously towards the protagonist directional indicator targets
(see \cite{turniness}),
while a novice player might think they must step with both feet at once, one for each arrow,
experts often find it more convenient (i.e., less overall foot motion) to use whichever single foot is closer at the time
to hit both arrows by
triggering one arrow's sensor with the heel and the other with the little bitty toesies.
Pads are typically constructed with a small triangular metal bracket at the corner of each arrow panel,
as shown in \Cref{fig:bracker-detail},
which the bridge of the foot must cross to achieve this, hence the name ``bracket jump''.
In case my prose explanation is not up to snuff, I also show in \Cref{fig:how-2-bracket}
a high-quality graphics render of a player's typical foot positions during a down+right bracket jump (henceforth ``DR'', et cetera).

@@

\begin{figure}[h]
	\begin{center}
		\includegraphics[width=0.48\textwidth]{how-2-bracket.jpg}
	\end{center}
	\caption{Down+right bracket jump real-world example.}
	\label{fig:how-2-bracket}
\end{figure}

@@

\section{Analyzing Bracket-Jumpiness}
\label{sec:analyzing}

@@

As before, the code is available at \url{https://github.com/bblum/sigbovik/blob/master/itg/code/ITG.hs}.
To give a sense of how much its elegance has been despoiled % hi william
since the last version:
% TODO: figure out these numbers without counting comments
98 lines of Haskell (which computed crossovers, footswitches, jacks, doublesteps, \textit{and} crossover footswitches)
has grown to 248 lines (just to handle this one. new. feature),
and the once-simple datatype definition of

@@

In lieu of an algorithm listing, I'll simply show some of those most notable test cases in pictures and explain in prose how I handle them.

\subsection{Basic Examples}

Figure \Cref{fig:bracket-example}(a) shows a simple example where every jump (blue)
can be bracketed by the right foot,
alternating with

\subsection{Wackier Examples}

@@

I pose two evaluation questions, to be answered in the following subsections respectively.
\begin{enumerate}
	\item Which songs, packs, and/or step-cartographers are bracket jumpiest?
	\item Is bracket jumps the new crossovers or something?
\end{enumerate}

\subsection{Bracket Jumpiness}
\label{sec:bjness}

I measured the overal bracketiness of each chart in two ways,
first,
by comparing the number of bracket jumps against the total number of steps,
or the \textit{bracket jump step percentage} (BJS\%);
second,
by comparing against the total number of jumps only,
or the \textit{bracket jump jump percentage} (BJJ\%).
The BJS\% indicates a chart's overall density of steps requiring the player to step across the brackets
(which depends on how jumpy the chart is to begin with),
whereas the BJJ\% measures perhaps the author's intentionality in patterning their jumps either as brackets,
or as random where just some of them happen to be bracketable.

\Cref{tab:bjs} shows the leaderboard for the former metric, and \Cref{tab:bjj} the latter.%
\footnote{When measuring BJJ\% I excluded charts with fewer than 10 jumps in total,
which put up a handful of false positives in the 100\% ranks.}
Note that the aforementioned recent Technical Showcase pack series put up three bracketiest charts in the BJJ\% category;
meanwhile, the UPS packs, known for their gimmicks and general lack of respect for player comfort,
secure several top spots on the BJS\% board.
For comparison's sake, the overall BJJ\% of the entire corpus is 20\% (93k/470k),
meaning that roughly 1 in 5 randomly-patterned jumps are by chance bracketable.
The overall BJS\% (a thoroughly less meaningful statistic) is 0.7\%.

% TODO check margins
\begin{table}[t]
	\begin{center}
		\small
	\begin{tabular}{r|l|l|r|r}
		\bf Ft. & \bf Name & \bf Pack & \bf \#BJ & \bf BJS\% \\
		\hline
		10 & SOBA			& Squeaky Beds \&c.	& 366 & 83\% \\
		11 & Get Off of My Way		& UPS 3			& 115 & 28\% \\
		12 & Hardware Store		& Keyboard Coll. III	& 207 & 23\% \\
		13 & Firestorm			& BemaniBeats 3		& 153 & 31\% \\
		14 & Bounce			& UPS 2			& 126 & 21\% \\
		15 & Ikaros Dynamite!!!!	& UPS 2			& 238 & 34\% \\
		16 & Mermaid Island		& Tachyon Alpha		& 346 & 49\% \\
		17 & Toccata \& Fugue		& CuoReNeRo M.P.	& 255 & 24\% \\
		% this chart is ALL jumps so it's kinda false positive
		%420 & Paa			& UPS 4			& 263 & 34\% \\
	\end{tabular}
	\end{center}
	\caption{Charts of each difficulty with the highest bracket jump density among all steps (BJS\%).}
	\label{tab:bjs}
\end{table}
\begin{table}[t]
	\begin{center}
		\small
	\begin{tabular}{r|l|l|r|r}
		\bf Ft. & \bf Name & \bf Pack & \bf \#BJ & \bf BJJ\% \\
		\hline
		 9 & Drifting Away		& UPS 4			&  26 &  88\% \\
		10 & SOBA			& Squeaky Beds \&c.	& 366 &  99\% \\
		11 & Save Miracles		& ECFA 2019		&  42 &  95\% \\
		12 & Electrical Paradise	& Chic. Timing Auth.	&  35 & 100\% \\
		12 & Encore			& Tech. Showcase	&  29 & 100\% \\
		12 & Nemeton			& Subluminal		&  12 & 100\% \\
		13 & Decadent Dandy		& Tech. Showcase 3	&  12 & 100\% \\
		14 & Nageki no Ki		& Valex's M.4-A.A. 8	& 116 &  94\% \\
		15 & Beach Party		& Tech. Showcase 2	& 162 &  80\% \\
		16 & Mermaid Island		& Tachyon Alpha		& 346 &  79\% \\
		17 & Zombie Sunset		& Jummy Jawns 2		& 325 &  88\% \\
		69 & koopa bling		& UPS 4			&  39 &  98\% \\
	\end{tabular}
	\end{center}
	\caption{Charts of each difficulty with the highest percentage of their total jumps bracketable (BJJ\%).
	All participants of the 3-way tie for 12-footers are listed.}
	%filtered by having at least 10 jumps to begin with.}
	\label{tab:bjj}
\end{table}

Next up is bracket jumpiness by pack and by author.
\Cref{tab:pack-bjs,tab:pack-bjj} show the top 10 packs in both bracketiness metrics,
and \Cref{tab:author-bjs,tab:author-bjj} the top 10 authors
(filtered by having written at least 10 charts).
For the packs, I also list the release year of each;
note how despite the relatively even spread of years in BJS\%,
2018 dominates the BJJ\% leaderboard,
suggesting that older charts' bracket jumps arose by chance simply from having a lot of jumps to begin with.
I'll come back to this point in \Cref{sec:eval-years}.
In honorable mentionth place is chart author Halogen-, whose 11 charts contain 154 jumps, \textit{none} of which are bracketable.
What a purist!

@@

\begin{table}[t]
	\begin{center}
		\small
	\begin{tabular}{l|r|r}
		\bf Author name & \bf \#charts & \bf BJS\% \\
		\hline
		Paul J Kim & 45 & 9.9 \\
		sssmsm & 45 & 4.4 \\
		Liam & 11 & 3.4 \\
		Snooze & 22 & 2.5 \\
		M. Emirzian & 26 & 2.2 \\
		bblum & 40 & 1.9 \\
		Ninevolt & 10 & 1.7 \\
		B. Vergara & 95 & 1.6 \\
		C. Emirzian & 18 & 1.6 \\
		B. Dinh & 10 & 1.6 \\
		Renard & 51 & 1.5 \\
	\end{tabular}
	\end{center}
	\caption{Step cartographers who write charts densest in bracket jumps.}
	\label{tab:author-bjs}
\end{table}

@@

I sought to prove \cite{dril}'s claim that bracket jumps have skyrocketed in popularity in the most recent years of stepchart-making,
%I decided to do this
by finding the progression of bracketiness across past years of stepchart packs.
However,
actually assigning a firm release date to every pack on my hard drive turned out to be a feat of internet archaeology unto itself,
even involving archive.org for one step.
I'll spare you the details, but you can peruse them in the second spreadsheet of the dataset linked at the start of this section.
Ultimately, all packs older than 2013 had to be grouped together,
as no pack collection website, facebook group, or laptop filesystem could accurately date enough packs
before then to give meaningful sample sizes for each year.
I also decided to exclude tournament packs such as ECFA from this analysis,
which curate existing stepcharts written possibly long ago.

@@

\begin{figure}[t]
	\begin{tabular}{c}
		\includegraphics[width=0.42\textwidth]{step_plot.pdf}
		\\
		(a) BJS\% across years ($\beta$ = 0.085\%, CI = \{0.032,0.14\}).
		\\
		\\
		\includegraphics[width=0.42\textwidth]{jump_plot.pdf}
		\\
		(a) BJJ\% across years ($\beta$ = 1.30\%, CI = \{0.40,2.21\}).
	\end{tabular}
	\caption{Linear best fit of bracket jumps' popularity across the years of ITG's history.
	Grey area depicts 95\% CI.}
	\label{fig:stats}
\end{figure}

\textit{Bias.}
This analysis is prone to selection bias in my own pack downloading habits:
supposing I suddenly became more interested in playing brackety charts recently,
that would certainly influence how many bracket jumps appeared each year in this corpus.
However, I believe the BJJ\% measurement adequately compensates for this,
for if I suddenly became interested in stamina instead of technical and added a bunch of 10-minute trance packs,
that would impact only BJS\%.
As far as I know, I show no preference for more or less jumpy charts overall than the community average,
leading me to trust the BJJ\% test.

@@

% TODO: check figure placement
\begin{figure*}[t]
	\begin{center}
	\begin{tabular}{c}
		\includegraphics[width=0.16\textwidth]{disco-beachparty.png}
		\\
		(a) Beach Party, \\ 15 \cite{beachparty} %\\
	\end{tabular}
	\begin{tabular}{c}
		\includegraphics[width=0.16\textwidth]{disco-koopa.png}
		\\
		(b) koopa bling, \\ 69 \cite{koopa} %\\
	\end{tabular}
	\begin{tabular}{c}
		\includegraphics[width=0.16\textwidth]{disco-zombiesunet.png}
		\\
		(c) Zombie Sunset, \\ 17 \cite{zombie} %\\
	\end{tabular}
	\begin{tabular}{c}
		\includegraphics[width=0.16\textwidth]{disco-owa.png}
		\\
		(d) 1-Winged Angel, \\ 14 \cite{owa} %\\
	\end{tabular}
	\begin{tabular}{c}
		\includegraphics[width=0.16\textwidth]{disco-paradise.png}
		\\
		(e) Electrical Paradise, \\ 12 \cite{e-paradise} %\\
	\end{tabular}
	\end{center}
	%\caption{Why I gotta write a overall caption, just read the prose on this one.}
	\caption{Example real-world charts with
	(a) false negative jumps that mines render bracketable after all,
	(b) jumps whose footing is affected by preceding holds,
	%false negative jumps forced to bracket by preceding holds, % XXX: this is a lie!! bonus pts if anyone spots it. i don't actually know why the alg didn't classify
	(c) brackets the chart author was sorry about,
	(d) ``retro'' style unbracketable jump-stream, %which, largely, cannot be bracketed,
	and
	(e) very intentionally many true bracket jumps from the BJJ\% category winner.}
	\label{fig:discussion}
\end{figure*}

@@

Many old charts written without regard to jump patterning allow for many of their jumps to be bracketed anyway.
I noticed an old stepchart for One-Winged Angel \cite{owa} ranking high in the list of overall total bracket jumps,
with 182,
simply because it's long and has a lot of steps,
but upon playing, the patterns definitely do not feel intentional,
and it even includes a section of stream with wholly unbracketable jumps interspersed.
\Cref{fig:discussion}(d) shows this section;
in this stream, the player would expect to step the red and blue arrows with her right foot and the lime ones with her left,
but note how the jumps (hint: all in blue arrows, unless you're reading this on dead trees)
occur as a mix of left and right arrows, with even a UD ``candle'' jump making an appearance.
However, it measures an unremarkable 9 BJS\% and 38 BJJ\%---I just wanted to feature it
as an example of what non-bracketable jump stream looks like.

Authors seem to realize when their charts are too bracket jumpy for comfort:
several of the charts appearing in the highest ranks of the spreadsheet had their step author field filled in as ``Stupid'' or ``I'm sorry''.
In fact, I personally found the charts with 100 BJJ\%, i.e., \textit{all} jumps appearing therein were bracketable,
to be more tasteful than those with more total BJS\% but a few normal jumps as well.
I'd hypothesize this is because to achieve 100 BJJ\% requires a certain intentionality,
resulting in better chart design overall
(note that such charts necessarily include no LR or DU jumps whatsoever).
Either that, or the more uncomfortable ones include certain extremely turny/candley/doublesteppy patterns
around their brackets that causes the algorithm to count them as normal jumps instead.
Anyway, \Cref{fig:discussion}(e) shows the chart with most bracket jumps (35) among ones with 100 BJJ\%,
which I tried out for myself and found quite enjoyable \cite{e-paradise}.
Amusingly, Encore \cite{encore}
features a Hard chart in addition to its Challenge,
which former ranks second place to Electrical Paradise among 100 BJJ\%s with 29 brackets.
Upon inspection, I found this chart identical to the Challenge (upcoming in \Cref{fig:xb}(a)),
only the crossover brackets
(to be discussed later in \Cref{sec:xover-brackers})
having been replaced by easier, normal ones.

@@

However, ambitious stepchart authors have recently experimented with encouraging the player to bracket jumps while crossed-over.
Such charts attempt to force, or at least hint, these jumps to be bracketed
via additional stepchart elements: in \Cref{fig:xb}(a), mines on the left arrow force the player to remove her left foot in preparation for the crossover (and being in the middle of fast stream, further discourages doublestepping),
and in \Cref{fig:xb}(b), extending one of the arrows as a hold encourages the player to bracket the subsequent jump with the other foot.
In (b)'s case, the player must also switch feet on each of the non-hold arrows,
effectively making these crossover-footswitch-brackets,
which would truly be a wonder to identify programmatically.
% these can be cheated anyway
Note however the high measure counter in both, meaning these occur quite late in the song;
both charts first spend some time ``teaching'' the player both to expect bracket jumps
and how it intends to signal crossovers in general---so if a human is not expected to understand them at first glance,
I think it is fair to leave out of the algorithm too for now.

\begin{figure}[t]
	\begin{center}
	\begin{tabular}{cc}
	\begin{tabular}{c}
		\includegraphics[width=0.16\textwidth]{xb-encore.png}
		\\
		(a) Encore, 12 \\ \cite{encore} %\\
		%(12, Tech. Showcase 1)
	\end{tabular}
		&
	\begin{tabular}{c}
		\includegraphics[width=0.16\textwidth]{xb-atariwave.png}
		\\
		(a) Atariwave, 10 \\ \cite{atariwave} %\\
		%(10, Tech. Showcase 3)
	\end{tabular}
	\end{tabular}
	\end{center}
	\caption{Crossover bracket jumps must be signaled via either (a) mines or (b) holds to remove ambiguity.}
	\label{fig:xb}
\end{figure}

@@

@@

% TODO: is it true.
Bracket jumps is, in fact, the new crossovers or something.
%I never know what to put down here. Much love to the community..!

@@

neuropantser (Michael T. Lawson) provided invaluable eleventh-hour (literally) advice
on \Cref{sec:eval-years}'s statistical analysis, up to and including making the graphs for me.
And much love to the SIGBOVIK community.

>>>>

What, then, is a bracket-jump (which I shall \textit{not}, henceforth, abbreviate for brevity)?
Put simply, whenever two arrow-shaped obstacles proceed simultaneously towards the protagonist directional indicator targets
(see \cite{turniness}),
while a novice player might think they must step with both feet at once, one for each arrow,
experts often find it more convenient (i.e., less overall foot motion) to use whichever single foot is closer at the time
to hit both arrows by
triggering one arrow's sensor with the heel and the other with the little bitty toesies.
Pads are typically constructed with a small triangular metal bracket at the corner of each arrow panel,
as shown in \Cref{fig:bracker-detail},
which the bridge of the foot must cross to achieve this, hence the name ``bracket jump''.
In case my prose explanation is not up to snuff, I also show in \Cref{fig:how-2-bracket}
a high-quality graphics render of a player's typical foot positions during a down+right bracket-jump (henceforth ``DR'', et cetera).

@@

\begin{figure}[h]
	\begin{center}
		\includegraphics[width=0.3625\textwidth]{how-2-bracket.jpg}
	\end{center}
	\caption{Down+right bracket-jump real-world example.}
	\label{fig:how-2-bracket}
\end{figure}

@@

\section{Algorithm Design}
\label{sec:analyzing}

@@

As before, the code is available at \url{https://github.com/bblum/sigbovik/blob/master/itg/code/ITG.hs}.
To give a sense of how much its elegance has been despoiled % hi william
since the last version:
62 lines of Haskell (which computed crossovers, footswitches, jacks, doublesteps, \textit{and} crossover footswitches)
has grown to 156 lines (just to handle this one. new. feature),
and the once-simple datatype definition of

@@

The general gist is that in addition to tracking which foot stepped the last arrow and optimizing for alternating feet,
we also track which arrow(s) each foot was last on,
and whenever we encounter a LD/LU/DR/UR jump, bracket it if both:
\begin{enumerate}
	\item The last foot is opposite the foot required for the jump (e.g. L for LD), and
	\item The last foot's last note(s) does not intersect the arrows involved in the jump (i.e., R not on D preceding LD).
\end{enumerate}
In lieu of a detailed algorithm listing, I'll simply show some of those most notable test cases in pictures and explain in prose how the implementation must handle them,
which was pretty much my implementation strategy to begin with.

\Cref{fig:tests} shows a bunch of patterns with a bracketable jump,
and each subfigure in \Cref{fig:tests2} shows a corresponding unbracketable case.
%with a similar pattern.
Whenever two jumps are pictured in the same unit test, we are concerned with whether the latter one is bracketable.
I'll now explain how the algorithm handles each case.

\begin{figure}[t]
	\begin{center}
	\begin{tabular}{ccc}
		\includegraphics[width=0.14\textwidth]{test-basic-yes.png} & 
		\includegraphics[width=0.14\textwidth]{test-footswitch-bracket-yes.png} &
		\includegraphics[width=0.14\textwidth]{test-reset-footing-yes.png} \\
		(a) All 4 OK
		&
		(b) LD, UR OK
		&
		(c) UR OK
		\\
		\\
		\includegraphics[width=0.14\textwidth]{test-propagate-through-ud-yes.png} &
		\includegraphics[width=0.14\textwidth]{test-reset-subsequent-stream-yes.png} &
		\includegraphics[width=0.14\textwidth]{test-footswitch-jack-yes.png} \\
		(d) LD OK
		&
		(e) UR, LU OK
		&
		(f) DR, LD OK
	\end{tabular}
	\end{center}
	\caption{Examples of bracketable jumps.}
	\label{fig:tests}
\end{figure}
\begin{figure}[t]
	\begin{center}
	\begin{tabular}{ccc}
		\includegraphics[width=0.14\textwidth]{test-basic-no.png} &
		\includegraphics[width=0.14\textwidth]{test-footswitch-bracket-no.png} &
		\includegraphics[width=0.14\textwidth]{test-reset-footing-no.png} \\
		(a) 2nd, 4th NG
		&
		(b) DR NG
		&
		(c) UR NG
		\\
		\\
		\includegraphics[width=0.14\textwidth]{test-propagate-through-ud-no.png} &
		\includegraphics[width=0.14\textwidth]{test-reset-subsequent-stream-no.png} &
		\includegraphics[width=0.14\textwidth]{test-footswitch-jack-also-yes.png} \\
		(d) LU NG
		&
		(e) 2nd UR NG
		&
		(f) Trick question
	\end{tabular}
	\end{center}
	\caption{Examples of unbracketable jumps,
	each corresponding to the similar pattern in \Cref{fig:tests}.}
	\label{fig:tests2}
\end{figure}

\renewcommand{\theenumi}{\alph{enumi}}
\begin{enumerate}
	\item In the OK case, the right foot can bracket every jump (condition 1),
		as the left foot is always on a different arrow than the jump (condition 2).
		In the NG case, condition 2 is violated for the second and fourth jumps.
	\item Version of (a) testing that preceding bracket-jumps also contribute to condition 2.
		The NG case would be a footswitch bracket requiring rather uncomfortable footing angles.
	\item In OK, LR cannot be bracketed, but clears the prior footings so UR can.
		In NG, LR can be bracketed, meaning UR cannot while still alternating feet (condition 1).
	\item DU jumps can never be bracketed, but can be stepped facing either direction, and this must respect prior footings.
		In both cases DU is stepped with right foot U, which allows bracketing LD but not LU thereafter.
	\item In OK, the crossover does not interfere with either bracket and the player can alternate feet throughout.
		In NG, either doublestepping or jumping normally is required
		(or a future-work crossover bracket; see \Cref{sec:xover-brackers}).
	\item In OK, the player easily footswitches on U.
		In the corresponding case, the player can actually just jack U instead, stepping it twice with her left foot
		before bracketing DR again with her right.
\end{enumerate}
\renewcommand{\theenumi}{\arabic{enumi}}

These examples are available as {\tt test-bracket-*.sm} unit tests in the code repository.
Hopefully armed with this understanding of the bracket jumping rules, let's now move on to real-world stepcharts.

% non bracket jumps reset the lastarrowLR

% no footswitch bracket - angles of yr ankles hurt
% propagate thru UD
% tiebreak preceding stream
% retroactively cancel if last stream hadta switch

@@

I pose two evaluation questions, to be answered in the following subsections respectively.
\begin{enumerate}
	\item Which songs, packs, and/or step-cartographers are bracket-jumpiest?
	\item Is bracket jumps the new crossovers or something?
\end{enumerate}

\subsection{Bracket-Jumpiness}
\label{sec:bjness}

I measured the overal bracketiness of each chart in two ways,
first,
by comparing the number of bracket jumps against the total number of steps,
or the \textit{bracket-jump step percentage} (BJS\%);
second,
by comparing against the total number of jumps only,
or the \textit{bracket jump jump percentage} (BJJ\%).
The BJS\% indicates a chart's overall density of steps requiring the player to step across the brackets
(which depends on how jumpy the chart is to begin with),
whereas the BJJ\% measures perhaps the author's intentionality in patterning their jumps either as brackets,
or as random where just some of them happen to be bracketable.

\Cref{tab:bjs} shows the leaderboard for the former metric, and \Cref{tab:bjj} the latter.%
\footnote{When measuring BJJ\% I excluded charts with fewer than 10 jumps in total,
which put up a handful of false positives in the 100\% ranks.}
Note that the aforementioned recent Technical Showcase pack series put up three bracketiest charts in the BJJ\% category;
meanwhile, the UPS packs, known for their gimmicks and general lack of respect for player comfort,
secure several top spots on the BJS\% board.
For comparison's sake, the overall BJJ\% of the entire corpus is 19.8\% (93k/470k),
meaning that roughly 1 in 5 randomly-patterned jumps are by chance bracketable.
The overall BJS\% (a thoroughly less meaningful statistic) is 0.7\%.

\begin{table}[t]
	\begin{center}
		\small
	\begin{tabular}{r|l|l|r|r}
		\bf Ft. & \bf Name & \bf Pack & \bf \#BJ & \bf BJS\% \\
		\hline
		10 & SOBA			& Squeaky Beds \&c.	& 366 & 83\% \\
		11 & Get Off of My Way		& UPS 3			& 115 & 28\% \\
		12 & Hardware Store		& Keyboard Coll. III	& 207 & 23\% \\
		13 & Firestorm			& BemaniBeats 3		& 153 & 31\% \\
		14 & Bounce			& UPS 2			& 126 & 21\% \\
		15 & Ikaros Dynamite!!!!	& UPS 2			& 238 & 34\% \\
		16 & Mermaid Island		& Tachyon Alpha		& 346 & 49\% \\
		17 & Toccata \& Fugue		& CuoReNeRo M.P.	& 255 & 24\% \\
		% this chart is ALL jumps so it's kinda false positive
		%420 & Paa			& UPS 4			& 263 & 34\% \\
	\end{tabular}
	\end{center}
	\caption{Charts of each difficulty with the highest bracket-jump density among all steps (BJS\%).}
	\label{tab:bjs}
\end{table}
\begin{table}[t]
	\begin{center}
		\small
	\begin{tabular}{r|l|l|r|r}
		\bf Ft. & \bf Name & \bf Pack & \bf \#BJ & \bf BJJ\% \\
		\hline
		 9 & Drifting Away		& UPS 4			&  26 &  88\% \\
		10 & SOBA			& Squeaky Beds \&c.	& 366 &  99\% \\
		11 & Save Miracles		& ECFA 2019		&  42 &  95\% \\
		12 & Electrical Paradise	& Chic. Timing Auth.	&  35 & 100\% \\
		12 & Encore			& Tech. Showcase	&  29 & 100\% \\
		12 & Nemeton			& Subluminal		&  12 & 100\% \\
		13 & Decadent Dandy		& Tech. Showcase 3	&  12 & 100\% \\
		14 & Nageki no Ki		& Valex's M.4-A.A. 8	& 116 &  94\% \\
		15 & Beach Party		& Tech. Showcase 2	& 162 &  80\% \\
		16 & Mermaid Island		& Tachyon Alpha		& 346 &  79\% \\
		17 & Zombie Sunset		& Jummy Jawns 2		& 325 &  88\% \\
		69 & koopa bling		& UPS 4			&  39 &  98\% \\
	\end{tabular}
	\end{center}
	\caption{Charts of each difficulty with the highest percentage of their total jumps bracketable (BJJ\%).
	All participants of the 3-way tie for 12-footers are listed.}
	%filtered by having at least 10 jumps to begin with.}
	\label{tab:bjj}
\end{table}

Next up is bracket jumpiness by pack and by author.
\Cref{tab:pack-bjs,tab:pack-bjj} show the top 10 packs in both bracketiness metrics,
and \Cref{tab:author-bjs,tab:author-bjj} the top 10 authors
(filtered by having written at least 10 charts).
For the packs, I also list the release year of each;
note how despite the relatively even spread of years in BJS\%,
2018 dominates the BJJ\% leaderboard,
suggesting that older charts' bracket-jumps arose by chance simply from having a lot of jumps to begin with.
I'll come back to this point in \Cref{sec:eval-years}.
In honorable mentionth place is chart author Halogen-, whose 11 charts contain 154 jumps, \textit{none} of which are bracketable.
What a purist!

@@

I sought to prove \cite{dril}'s claim that bracket-jumps have skyrocketed in popularity in the most recent years of stepchart-making,
%I decided to do this
by finding the progression of bracketiness across past years of stepchart packs.
However,
actually assigning a firm release date to every pack on my hard drive turned out to be a feat of internet archaeology unto itself,
even involving archive.org for one step.
I'll spare you the details, but you can peruse them in the second spreadsheet of the dataset linked at the start of this section.
Ultimately, all packs older than 2013 had to be grouped together,
as no pack collection website, facebook group, or laptop filesystem could accurately date enough packs
before then to give meaningful sample sizes for each year.
I also decided to exclude tournament packs such as ECFA from this analysis,
which curate existing stepcharts written possibly long ago.

\begin{table}[t]
	\begin{center}
		\small
	\begin{tabular}{l|r|r}
		\bf Author name & \bf \#charts & \bf BJS\% \\
		\hline
		Paul J Kim & 45 & 9.9 \\
		sssmsm & 45 & 4.4 \\
		Liam & 11 & 3.4 \\
		Snooze & 22 & 2.5 \\
		M. Emirzian & 26 & 2.2 \\
		bblum & 40 & 1.9 \\
		Ninevolt & 10 & 1.7 \\
		B. Vergara & 95 & 1.6 \\
		C. Emirzian & 18 & 1.6 \\
		B. Dinh & 10 & 1.6 \\
		Renard & 51 & 1.5 \\
	\end{tabular}
	\end{center}
	\caption{Step cartographers who write charts densest in bracket-jumps.}
	\label{tab:author-bjs}
\end{table}

@@

@@

\begin{figure}[t]
	\begin{tabular}{c}
		\includegraphics[width=0.42\textwidth]{step_plot.pdf}
		\\
		(a) BJS\% across years ($\beta$ = 0.085\%/yr, CI = \{0.032,0.14\}).
		\\
		\\
		\includegraphics[width=0.42\textwidth]{jump_plot.pdf}
		\\
		(a) BJJ\% across years ($\beta$ = 1.30\%/yr, CI = \{0.40,2.21\}).
	\end{tabular}
	\caption{Linear best fit of bracket jumps' popularity across the years of ITG's history.
	Grey area depicts 95\% CI.}
	\label{fig:stats}
\end{figure}

\textit{Bias.}
This analysis is prone to selection bias in my own pack downloading habits:
supposing I suddenly became more interested in playing brackety charts recently,
that would certainly influence how many bracket-jumps appeared each year in this corpus.
However, I believe the BJJ\% measurement adequately compensates for this,
for if I suddenly became interested in stamina instead of technical and added a bunch of 10-minute trance packs,
that would impact only BJS\%.
As far as I know, I show no preference for more or less jumpy charts overall than the community average,
leading me to trust the BJJ\% test.

@@

% TODO: check figure placement
\begin{figure*}[t]
	\begin{center}
	\begin{tabular}{c}
		\includegraphics[width=0.16\textwidth]{disco-beachparty.png}
		\\
		(a) Beach Party, \\ 15 \cite{beachparty} %\\
	\end{tabular}
	\begin{tabular}{c}
		\includegraphics[width=0.16\textwidth]{disco-koopa.png}
		\\
		(b) koopa bling, \\ 69 \cite{koopa} %\\
	\end{tabular}
	\begin{tabular}{c}
		\includegraphics[width=0.16\textwidth]{disco-zombiesunet.png}
		\\
		(c) Zombie Sunset, \\ 17 \cite{zombie} %\\
	\end{tabular}
	\begin{tabular}{c}
		\includegraphics[width=0.16\textwidth]{disco-owa.png}
		\\
		(d) 1-Winged Angel, \\ 14 \cite{owa} %\\
	\end{tabular}
	\begin{tabular}{c}
		\includegraphics[width=0.16\textwidth]{disco-paradise.png}
		\\
		(e) Electrical Paradise, \\ 12 \cite{e-paradise} %\\
	\end{tabular}
	\end{center}
	%\caption{Why I gotta write a overall caption, just read the prose on this one.}
	\caption{Example real-world charts with
	(a) false negative jumps that mines render bracketable after all,
	(b) jumps whose footing is affected by preceding holds fixing one foot in place,
	%false negative jumps forced to bracket by preceding holds, % XXX: this is a lie!! bonus pts if anyone spots it. i don't actually know why the alg didn't classify
	(c) brackets the chart author was sorry about,
	(d) ``retro'' style unbracketable jump-stream, %which, largely, cannot be bracketed,
	and
	(e) intentionally many true bracket jumps from the BJJ\% category winner.}
	\label{fig:discussion}
\end{figure*}

@@

Many old charts written without regard to jump patterning allow for many of their jumps to be bracketed anyway.
I noticed an old stepchart for One-Winged Angel \cite{owa} ranking high in the list of overall total bracket-jumps,
with 182,
simply because it's long and has a lot of steps,
but upon playing, the patterns definitely do not feel intentional,
and it even includes a section of stream with wholly unbracketable jumps interspersed.
\Cref{fig:discussion}(d) shows this section;
in this stream, the player would expect to step the red and blue arrows with her right foot and the lime ones with her left,
but note how the jumps (hint: all in blue arrows, unless you're reading this on dead trees)
occur as a mix of left and right arrows, with even a UD ``candle'' jump making an appearance.
However, it measures an unremarkable 9 BJS\% and 38 BJJ\%---I just wanted to feature it
as an example of what non-bracketable jump stream looks like.

Authors seem to realize when their charts are too bracket jumpy for comfort:
several of the charts appearing in the highest ranks of the spreadsheet had their step author field filled in as ``Stupid'' or ``I'm sorry''.
In fact, I personally found the charts with 100 BJJ\%, i.e., \textit{all} jumps appearing therein were bracketable,
to be more tasteful than those with more total BJS\% but a few normal jumps as well.
I'd hypothesize this is because to achieve 100 BJJ\% requires a certain intentionality,
resulting in better chart design overall
(note that such charts necessarily include no LR or DU jumps whatsoever).
Either that, or the more uncomfortable ones include certain extremely turny/candley/doublesteppy patterns
around their brackets that causes the algorithm to count them as normal jumps instead.
Anyway, \Cref{fig:discussion}(e) shows the chart with most bracket-jumps (35) among ones with 100 BJJ\%,
which I tried out for myself and found quite enjoyable \cite{e-paradise}.
Amusingly, Encore \cite{encore}
features a Hard chart in addition to its Challenge,
which former ranks second place to Electrical Paradise among 100 BJJ\%s with 29 brackets.
Upon inspection, I found this chart identical to the Challenge (upcoming in \Cref{fig:xb}(a)),
only the crossover brackets
(to be discussed later in \Cref{sec:xover-brackers})
having been replaced by easier, normal ones.

@@

@@

However, ambitious stepchart authors have recently experimented with encouraging the player to bracket-jumps while crossed-over.
Such charts attempt to force, or at least hint, these jumps to be bracketed
via additional stepchart elements: in \Cref{fig:xb}(a), mines on the left arrow force the player to remove her left foot in preparation for the crossover (and being in the middle of fast stream, further discourages doublestepping),
and in \Cref{fig:xb}(b), extending one of the arrows as a hold encourages the player to bracket the subsequent jump with the other foot.
In (b)'s case, the player must also switch feet on each of the non-hold arrows,
effectively making these crossover-footswitch-brackets,
which would truly be a wonder to identify programmatically.
% these can be cheated anyway
Note however the high measure counter in both, meaning these occur quite late in the song;
both charts first spend some time ``teaching'' the player both to expect bracket jumps
and how it intends to signal crossovers in general---so if a human is not expected to understand them at first glance,
I think it is fair to leave out of the algorithm too for now.

\begin{figure}[t]
	\begin{center}
	\begin{tabular}{cc}
	\begin{tabular}{c}
		\includegraphics[width=0.16\textwidth]{xb-encore.png}
		\\
		(a) Encore, 12 \\ \cite{encore} %\\
		%(12, Tech. Showcase 1)
	\end{tabular}
		&
	\begin{tabular}{c}
		\includegraphics[width=0.16\textwidth]{xb-atariwave.png}
		\\
		(a) Atariwave, 10 \\ \cite{atariwave} %\\
		%(10, Tech. Showcase 3)
	\end{tabular}
	\end{tabular}
	\end{center}
	\caption{Crossover bracket-jumps must be signaled via either (a) mines or (b) holds to remove ambiguity.}
	\label{fig:xb}
\end{figure}

@@

Bracket jumps is, in fact, statistically significantly, the new crossovers or something.
%I never know what to put down here. Much love to the community..!

@@

neuropantser (Michael T. Lawson) provided invaluable eleventh-hour (literally) advice
on \Cref{sec:eval-years}'s statistical analysis, up to and including making the graphs for me.
Greg Hanneman unknowingly helped copy-edit my hyphenation.
And much love to the SIGBOVIK community.
